Modelling the travel time of transit vehicles in \rt through a GTFS-based road network using GPS vehicle locations

### 2019-08-03 anzjs2018/sections/04_implementation.tex
Fix graphs

\begin{figure}[tb]
    \centering
    \includegraphics[width=0.7\textwidth]{figures/04_model_results_timing.pdf}
    \caption{
        Application timings were recorded for each component of each
        iteration, and averages over a 15~minute window are shown here.
        The timings shown are \emph{wall time}, the actual time passed,
        not CPU time.
        Steps that only involve particle calculations are approximately
        of complexity $O(N)$,
        however some steps are a little more complex (writing ETAs involves computing quantiles), 
        so there is a slightly exponential trend.
    }
        % The timings for various parts of the application, and overall. %
        % The trend is slightly non-linear due to the complexity of the ETA generation step, %
        % which requires sorting of the particle ETAs to obtain quantile estimates.}
    \label{fig:timings}
\end{figure}

@@

\begin{figure}[tb]
    \centering
    \includegraphics[width=0.7\textwidth]{figures/04_model_results_dist.pdf}
    \caption{
        The distance between the observation and the nearest point on the shape
        is calculated by finding the shortest distance between point and path.
        This gives an approximate idea of GPS error, however it should be noted
        that many observations are made automatically at way points,
        which accounts for the peak at 1m.
        Note also that these observations have been truncated at 20m.
    }
    \label{fig:dist_to_route}
\end{figure}

Figure~\ref{fig:neff} shows that $N_\text{eff}$ increased
with GPS error and decreased with system noise,
while remaining unaffected by changing $N$.
The degeneration rate decreased with GPS error and increased with $N$,
and was unaffected by system noise,
as seen in Figure~\ref{fig:degen_rate}.
Large GPS error affects the likelihood,
giving more weight to particles farther from the observation
than does a smaller GPS error.
Conversely, increasing system noise spreads out the particle cloud,
so fewer particles will be near the vehicle.
Thus, these results are not surprising, 
but show that the model is working as expected.

\begin{figure}[tb]
    \centering
    \includegraphics[width=0.9\textwidth]{figures/04_model_results_neff.pdf}
    \caption{
        The effective sample size is an approximate measure of the variance
        of the particle weights.
        When all particles have equal weights, $N_\text{eff} = N$,
        while if all the weight is on one particle, $N_\text{eff} = 1$. 
        The figure is subset by system noise, which shows how $N_\text{eff}$ 
        decreases as system noise increases,
        while increasing GPS error (on the $x$-axis) increases $N_\text{eff}$.
        Increasing $N$ has a negative effect on $N_\text{eff}$ but this is
        only evident when $N_\text{eff}$ is larger.
        The error bars show $\pm 2$SE.
    }
    \label{fig:neff}
\end{figure}

\begin{figure}[tb]
    \centering
    \includegraphics[width=0.9\textwidth]{figures/04_model_results_degen.pdf}
    \caption{
        The degeneration rate is the percentage of iterations where there were no
        plausible particles, so the vehicle was ``lost'' by the model.
        This decreases as GPS error and number of particles increase,
        but there is no visible effect of system noise.
        Note, however, that some of these can be attributed to invalid data,
        for example if the driver has entered the wrong trip direction.
        The error bars show $\pm 2$SE.
    }
    \label{fig:degen_rate}
\end{figure}

The last measure of performance is the relative variance of segment travel times.
We used relative variance as each segment has a very different travel time distribution,
so to compare the variability between simulations we needed to 
compute the average variance of travel times over all simulations,
taking the average (over road segments) of the ratio between variance for
each individual simulation and variance over all simulations.
Figure~\ref{fig:travel_times} shows that simulations with larger GPS error
resulted in a higher relative variance,
while simulations with more particles had lower relative variance.
There was no visible effect of system noise on the variability of travel times.

>>>>

\begin{figure}[tb]
    \centering
    \includegraphics[width=0.8\textwidth]{figures/04_model_results_timing.pdf}
    \caption{
        Application timings were recorded for each component of each
        iteration, and averages over a 15~minute window are shown here.
        The timings shown are \emph{wall time}, the actual time passed,
        not CPU time.
        Steps that only involve particle calculations are approximately
        of complexity $O(N)$,
        however some steps are a little more complex (writing ETAs involves computing quantiles), 
        so there is a slightly exponential trend.
    }
        % The timings for various parts of the application, and overall. %
        % The trend is slightly non-linear due to the complexity of the ETA generation step, %
        % which requires sorting of the particle ETAs to obtain quantile estimates.}
    \label{fig:timings}
\end{figure}

@@

\begin{figure}[tb]
    \centering
    \includegraphics[width=0.8\textwidth]{figures/04_model_results_dist.pdf}
    \caption{
        The distance between the observation and the nearest point on the shape
        is calculated by finding the shortest distance between point and path.
        This gives an approximate idea of GPS error, however it should be noted
        that many observations are made automatically at way points,
        which accounts for the peak at 1m.
        Note also that these observations have been truncated at 20m.
    }
    \label{fig:dist_to_route}
\end{figure}

Figure~\ref{fig:perf_stats} shows that $N_\text{eff}$ increased
with GPS error and decreased with system noise,
while remaining unaffected by changing $N$.
The degeneration rate decreased with GPS error and increased with $N$,
and was unaffected by system noise.
Large GPS error affects the likelihood,
giving more weight to particles farther from the observation
than does a smaller GPS error.
Conversely, increasing system noise spreads out the particle cloud,
so fewer particles will be near the vehicle.
Thus, these results are not surprising, 
but show that the model is working as expected.

\begin{figure}[tb]
    \centering
    \includegraphics[width=\textwidth]{figures/04_model_results_neff.pdf}
    \caption{
        The effective sample size is an approximate measure of the variance
        of the particle weights.
        When all particles have equal weights, $N_\text{eff} = N$,
        while if all the weight is on one particle, $N_\text{eff} = 1$. 
        The figure is subset by system noise, which shows how $N_\text{eff}$ 
        decreases as system noise increases,
        while increasing GPS error (on the $x$-axis) increases $N_\text{eff}$.
        Increasing $N$ has a negative effect on $N_\text{eff}$ but this is
        only evident when $N_\text{eff}$ is larger.
        The error bars show $\pm 2$SE.
    }
    \label{fig:perf_stats}
\end{figure}

% \begin{figure}[tb]
%     \centering
%     \includegraphics[width=0.9\textwidth]{figures/04_model_results_degen.pdf}
%     \caption{
%         The degeneration rate is the percentage of iterations where there were no
%         plausible particles, so the vehicle was ``lost'' by the model.
%         This decreases as GPS error and number of particles increase,
%         but there is no visible effect of system noise.
%         Note, however, that some of these can be attributed to invalid data,
%         for example if the driver has entered the wrong trip direction.
%         The error bars show $\pm 2$SE.
%     }
%     \label{fig:degen_rate}
% \end{figure}

The last measure of performance is the relative variance of segment travel times.
We used relative variance as each segment has a very different travel time distribution,
so to compare the variability between simulations we needed to 
compute the average variance of travel times over all simulations,
taking the average (over road segments) of the ratio between variance for
each individual simulation and variance over all simulations.
If $\tilde z_\ell^{\sigma,\epsilon}$ is a vector of all travel times
along segment $\ell$ during the simulation with GPS error $\sigma$ and system noise $\epsilon$,
then the relative variance $\bar v^{\sigma,\epsilon}$ is calculated by
\begin{equation*}
\begin{split}
v_\ell^{\sigma,\epsilon} &=
\frac{\mathrm{Var}(\tilde z_\ell^{\sigma,\epsilon})}{\mathrm{Var}(\tilde z_\ell)} \\
\bar v^{\sigma,\epsilon} &=
    \frac{1}{L}\sum_{\ell=1}^L v_\ell^{\sigma,\epsilon}
\end{split}
\end{equation*}
Figure~\ref{fig:travel_times} shows that simulations with larger GPS error
resulted in a higher relative variance,
while simulations with more particles had lower relative variance.
There was no visible effect of system noise on the variability of travel times.
